### 2022-06-13 sections/2-info-theory.tex
wip: new knn algorithm outline

The \textbf{distance function} must be a metric; that is it must have the properties that $\forall x, y, z \in X$:
\begin{enumerate}[i.]
    \item $f$ is deterministic.
    \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$.
    \item $f(x, y) = f(y, x)$, i.e. symmetry.
    \item $f(x, z) \leq f(x, y) + f(y, z)$, i.e. the triangle inequality.
\end{enumerate}

>>>>

The following analysis assumes that the \textbf{distance function} is a metric;
i.e. $\forall x, y, z \in X$:
\begin{enumerate}[i.]
    \item $f$ is deterministic.
    \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$, i.e. identity.
    \item $f(x, y) = f(y, x)$, i.e. symmetry.
    \item $f(x, z) \leq f(x, y) + f(y, z)$, i.e. the triangle inequality.
\end{enumerate}

See Section~\ref{sec:discussion} for a discussion on what happens with distance functions for which these properties do not hold.

### 2022-06-13 sections/3-methods.tex
A comment
wip: new knn algorithm outline

Modified binary search.
Lift this from CHESS paper.

@@

Repeated $\rho$-NN search with where radius is modified with each iteration as guided by lfd of query ball.

Lift from CHESS paper as a starting point.

Complexity is some factor times CHESS-search.

>>>>

Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.

Modified binary search to identify candidate leaf clusters.
Exhaustive leaf search over those leaves.

Lift this from CHESS paper. Be sure to cite CHESS here.

@@

Given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.

Given a Cluster $C$, let $c$ be its center and $r$ be its radius.
Let $\delta^0 = f(q, c)$ be the distance from the query to the cluster center.
Let $\delta^1 = \delta^0 + r$ be the distance from the query to the potentially farthest instance in the Cluster.
Let $\delta^2 = max(0, \delta^0 - r)$ be the distance from the query to the potentially closest instance in the Cluster.

\subsubsection{Tree Search}
\label{subsubsec:methods:knn-search:tree-search}

Let $candidates$ be a priority queue sorted by non-decreasing $\delta^1$.
Initialize $candidates$ with the root cluster.
Replace clusters in $candidates$ by their children until $candidates$ has at least $k$ clusters.

Take the $k^{th}$ candidate.
Any cluster with $\delta^2 \leq \delta^2_k$ could contain one of the $k$-nearest neighbors of $q$.
Keep all such clusters in $candidates$ and exclude all other clusters.
Expand search to the children of all $candidates$ so far and repeat until all $candidates$ are leaf clusters.

Sort, as a priority queue, $candidates$ by $\delta^2$ and break ties by $\delta^0$.
Proceed to Leaf Search with these candidates.

\subsubsection{Leaf Search}
\label{subsubsec:methods:knn-search:leaf-search}

Let $hits$ be a a fixed size priority queue with $k$ isntances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
Initialize $hits$ with the centers of $candidates$.

Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
Repeat until $candidates$ is empty.

\subsubsection{Complexity}
\label{subsubsec:methods:knn-search:complexity}

This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.
